\section{Situating LLM Advice: Platform Vernaculars and Aligned Judgment}
\label{sec:bg}

The central argument of this article is that LLMs can recognize harm without authorizing action and that this gap is systematic rather than incidental. Two ideas help specify that gap. \textit{Defamiliarization} names the methodological move: LLM divergence from community norms is treated as a probe that makes visible what habit and participation have rendered taken-for-granted within a platform's moral order. \textit{Vernacular competence} names the community-side capacity at stake: the locally ratified ability to convert recognition of harm into permission for action. The sections that follow situate this argument by specifying the moral order of the platform, the advisory defaults of LLMs, and the terms needed to interpret the gap between them.

\subsection{Platform Affordances and the Governance of Intimate Life}

Online advice communities do not simply host conversation; they produce normative orders \citep{gillespie2018custodians}. The upvote economies, comment threading, and anonymity structures of platforms like Reddit create platform vernaculars: community-specific ways of speaking, judging, and prescribing action that are inseparable from the infrastructures through which they circulate. On r/relationship\_advice, votes select for directiveness, moral certainty, and exit-orientation while rendering ambivalence invisible. The result is a concentrated mode of governing intimate life through distributed, informal mechanisms of judgment and sanction, with norms that differ markedly from professional counseling or earlier advice traditions \citep{reagle2025}.

The normative apparatus of r/relationship\_advice draws heavily on therapeutic vocabulary (gaslighting, narcissism, red flags, boundaries), exemplifying Illouz's ``therapeutic turn'' in which psychological categories become tools for distributing moral judgment \citep{illouz2008saving, rose1999governing}. On r/relationship\_advice, this vocabulary authorizes an unforgiving exit-orientation that treats relationship problems as puzzles with correct solutions. The community thus functions as a limiting case: a moral economy where voting has selected for maximal prescriptive clarity, producing norms concentrated enough that their absence in LLM outputs becomes measurable.

\subsection{LLMs and the Normative Defaults of Alignment}

Large language models encode the priorities and biases of their training data and alignment procedures \citep{bender2021dangers, santurkar2023opinions}. Because meaning in embedding spaces is constituted through distributional proximity across diverse sources \citep{brunila2025cosine}, the advice that emerges lacks the prescriptive force any specific community would recognize as adequate. Safety alignment through RLHF is one plausible contributor to this pattern, encouraging validation, hedging, and harm-avoidance as default advisory stances \citep{ouyang2022training}---shaping not just what models omit but the kinds of responses they often substitute in its place.

Research on LLM sycophancy documents this empirically. Cheng et al.'s ELEPHANT benchmark shows that models tend to preserve users' positive self-image and affirm both sides of moral conflicts \citep{cheng2025elephant_social_sycophancy}. In the domain of moral reasoning specifically, Sachdeva and van Nuenen find low inter-model agreement across everyday dilemmas, suggesting that variation in training and alignment is associated with a family of divergent moral stances, each internally consistent but collectively unmoored from any particular community's normative infrastructure \citep{Sachdeva2025}. These concerns intensify in advice contexts, where the stance a system takes---validating versus directing, balancing perspectives versus assigning responsibility---shapes what courses of action become thinkable for recipients \citep{vecchione2025artificial}.

\subsection{Vernacular Competence}

The concept that names what the community possesses and aligned models consistently forgo is \textit{vernacular competence}: the locally ratified capacity to convert recognition of harm into directive authorization for action. This is not a matter of vocabulary or tone alone. Top-voted comments routinely perform this conversion in a single move---``This is textbook abuse. Leave.’’ fuses an epistemic assessment with a deontic directive \citep{stevanovic2012deontic}. Through accumulated vote-selection, the community has developed a standard for what counts as advice in the full sense: utterances that not only interpret a situation but authorize a course of action with locally ratified force. Vernacular competence names that standard and the gap its absence produces.

LLM outputs make this standard visible by contrast. The comparison is methodologically useful because it defamiliarizes the community's own standards of adequacy \citep{shklovsky1917art, iliadis2016critical}: by introducing outputs calibrated to portable advisory defaults rather than subreddit-specific criteria, it renders visible what community participation ordinarily leaves unspoken. This works because LLMs are not simply another situated community with competing norms---they function as portable infrastructures with relatively stable advisory defaults across contexts. That stability is what makes the gap analytically productive rather than merely contrastive: it surfaces a recurring pattern in how these systems handle the move from harm recognition to action authorization.

The comparison also makes legible the community’s own taken-for-granted assumptions. On r/relationship\_advice, the conviction that controlling behavior warrants exit operates as a self-evident baseline, presupposing financial independence, material alternatives, and a therapeutic vocabulary through which exit is framed as self-care. These conditions do not need to be argued for because they are not experienced as conditions at all---they are simply what the community treats as obvious. Against that background, the models’ defaults become visible as a portable alternative, not a neutral one.