\section{Introduction: When Helpful AI Meets Platform Vernaculars}
\label{sec:intro}

A man posts to the Reddit community r/relationship\_advice about his girlfriend of three years. He has saved for eight years while living with his parents; she rents and earns less. He loves her, he says, but resents that the financial burden of buying a house will fall on him. He wonders whether to break up. The top-voted response does not validate his anxiety or explore his conflict: ``Honestly you should break up with her because she deserves someone better. You sound selfish... For you it's all about money.''

A large language model, given the same post, responds differently. It opens with gratitude (``Thank you for sharing your feelings so openly''), organizes his concerns into numbered considerations (``Acknowledge Your Feelings,'' ``Reflect on Your Priorities,'' ``Open Communication''), and closes by offering to ``help exploring specific options or ways to approach conversations with your girlfriend.'' It identifies his anxiety as understandable. It does not call him selfish.

As large language models become embedded in everyday advice settings \citep{vowels2024ai, yun2025online, huang2024chatgpt}, such encounters are increasingly common. Model responses enter domains already organized by their own moral orders---settings where what counts as good advice includes locally ratified ways of assigning blame, recognizing harm, and warranting action. The issue is less whether LLMs advise well in the abstract than how their aligned defaults---often shaped by corporate risk-mitigation and standardized safety protocols---interact with publics that have already developed situated standards for rendering judgment.

What is at stake here is a form of \textit{vernacular competence}: the community's locally ratified ability to move from recognizing harm to authorizing action, granting permission to exit, confront, or refuse in ways participants treat as adequate. Vernacular competence is not a universal standard but a situated achievement, relative to the norms a particular public develops and enforces.

r/relationship\_advice is not a proxy for human advice in general---it is a deliberately concentrated advisory formation whose platform affordances amplify prescriptive clarity. This, the current article argues, is precisely what makes divergence from it measurable. The discourse community's vote-ratified outputs make visible a normative infrastructure that would otherwise remain distributed and tacit, enabling a comparison that more diffuse settings cannot support.

The argument advanced here is not that aligned LLMs fail to reproduce this community’s locally ratified mode of judgment, or whether that judgment is `correct'. It is that models recognize many of the same dynamics the community does, yet systematically decline to convert that recognition into the forms of directive action the community treats as adequate. Models often deploy terms such as ``gaslighting'' or ``red flags'' at rates comparable to human commenters, but where the community converts diagnosis into directive action---leave, confront, refuse---models convert it into abstract entitlement (``you deserve respect'') or redirect toward reflection, communication, and therapeutic process, effectively neutralizing the friction of moral conflict. 

This pattern can be described as \textit{recognition without authorization}: the capacity to register harm while remaining reluctant to convert that recognition into directive permission for action. The gap is widest where it should, on an uncertainty account, be narrowest: on high-consensus posts involving abuse or safety threats, models recommend exit at roughly half the human rate. This pattern suggests that safety alignment might help shape everyday judgment not only through what models can say, but by standardizing the conditions under which harm is translated into warranted action.

Prior scholarship has documented pieces of this problem without assembling its structure. Research on LLM sycophancy and moral reasoning shows that models preserve users' positive self-image, affirm both sides of moral conflicts, and encode markedly different moral sensitivities across training pipelines \citep{sharma2024sycophancy, cheng2025elephant_social_sycophancy, 
Sachdeva2025}---but this work has largely treated divergence as a property of models rather than examining how aligned AI relates to the normative infrastructure that underlies community judgment, particularly in settings where that judgment is not encoded in discrete labels but in the accumulated, vote-ratified practice of converting diagnosis into warranted action.

Rather than treating divergence as error, this article treats it as a method for identifying the normative infrastructure of aligned AI. Because LLMs reproduce the vocabulary of relationship distress but not the scripts through which communities convert diagnosis into warranted action, they make visible not merely a local mismatch but a portable default, a standardized advisory style that validates, hedges, and redirects rather than authorizing action. Community norms make visible what alignment assumes as default: that therapeutic hedging is appropriate regardless of situational clarity. Model behavior, in turn, makes visible what the community assumes as obvious: that some situations call for directive judgment rather than open-ended reflection.

The article develops this argument through a comparison of four models' responses with community-endorsed advice across 11,565 r/relationship\_advice posts, using a high-consensus design that isolates cases where the community's moral certainty is strongest. The analysis makes three contributions. First, it documents a recurring pattern across four assistant-style models, which is described as \textit{recognition without authorization}. This pattern becomes visible when contrasted with the concentrated moral order of r/relationship\_advice. Second, it develops \textit{vernacular competence} as a concept for specifying what platform communities possess and aligned models often lack: not just knowledge of moral vocabulary, but situated norms for converting that vocabulary into warranted action. Third, it demonstrates that LLM divergence can serve as a systematic method for surfacing the tacit normative infrastructure of platform communities---an approach that could generalize beyond this single case. Taken together, the analysis reframes the encounter between AI and platform cultures as a site where the standardization of everyday moral judgment becomes visible.