% =============================================================================
% Title Page - Big Data & Society Submission
% =============================================================================
% This file contains all identifying information and will NOT be sent to reviewers.
% The main manuscript should be fully anonymized.
% =============================================================================

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{url}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}

\doublespacing

\begin{document}

\begin{center}
{\Large \textbf{TITLE PAGE}}\\[2em]
{\Large \textbf{(Not for peer review)}}
\end{center}

\vspace{2em}

% =============================================================================
% ARTICLE TITLE
% =============================================================================

\section*{Article Title}

Recognition Without Prescription: Defamiliarizing Relationship Advice Through LLM Divergence

% =============================================================================
% AUTHORS AND AFFILIATIONS
% =============================================================================

\section*{Authors}

\textbf{Tom van Nuenen}\\
D-Lab\\
University of California, Berkeley\\
Berkeley, CA, USA

% =============================================================================
% CORRESPONDING AUTHOR
% =============================================================================

\section*{Corresponding Author}

Tom van Nuenen\\
D-Lab\\
350 Social Sciences Building\\
University of California, Berkeley\\
Berkeley, CA 94720-3030\\
USA\\[1em]
Email: tomvannuenen@berkeley.edu\\
Phone: 5103659198

% =============================================================================
% ACKNOWLEDGMENTS
% =============================================================================

\section*{Acknowledgments}

Not applicable.

% =============================================================================
% STATEMENTS AND DECLARATIONS
% =============================================================================

\section*{Statements and Declarations}

\subsection*{Ethical Considerations}

This study analyzes publicly available data from Reddit (r/relationship\_advice). All posts analyzed were publicly accessible at the time of collection. No private or personally identifiable information was collected, and all quoted material in the manuscript has been paraphrased or presented in a manner that prevents identification of original posters. The research did not involve direct interaction with human subjects. Given the use of publicly available, anonymized data, formal ethics approval was not required for this study.

\subsection*{Consent to Participate}

Not applicable. This study analyzed publicly available online content and did not involve direct participation of human subjects.

\subsection*{Consent for Publication}

This study uses publicly available Reddit posts and comments. Usernames were not collected or reported, and quoted excerpts were selected from high-visibility posts to minimize the risk of identification. Because the data were publicly accessible and analyzed in aggregate, individual consent for publication was not required.

\subsection*{Declaration of Conflicting Interest}

The author declared no potential conflicts of interest with respect to the research, authorship, and/or publication of this article.

\subsection*{Funding}

The author received no financial support for the research, authorship, and/or publication of this article.

\subsection*{Data Availability}

The analysis code and processed data supporting this study are available at \url{https://anonymous.4open.science/r/recognition-without-prescription-7883}. Raw Reddit data were collected via the Pushshift API and are subject to Reddit's Terms of Service; the repository includes the post IDs necessary to reconstruct the dataset. 

\section*{Generative AI usage statement}

The author used Claude Opus 4.5 to assist with grammar checks, LaTeX formatting, and reviews of our code, as well as critiquing scripts and findings. All scientific contributions, analyses, interpretations, and final wording are the author's own, and the author takes full responsibility for the content.


% =============================================================================
% KEYWORDS
% =============================================================================

\section*{Keywords}

large language models; platform cultures; relationship advice; defamiliarization; vernacular competence; safety alignment; therapeutic governance; critical data studies

\end{document}
